---
title: "PA1_template"
author: "ZilongJi"
date: "Thursday, April 16, 2015"
output: html_document
---

Note: Before you run the .rmd, please make sure you are under the correct working directory

#### Part 1: LOADING AND PREPROCESSING THE DATA

1.Load the data(i.e. read.csv())
        
```{r echo=TRUE}
setwd("C:/Users/Administrator/Desktop/MLDM/Coursera/JHU data science/Reproducible research")
activity <- read.csv("activity.csv",header=TRUE,sep=",")
```

2.change the data frame to a data table

```{r}
library(data.table)
activity <- data.table(activity)
```

#### Part 2: WHAT IS MEAN TOTAL NUMBER OF STEPS TAKEN PER DAY?

1.Make a histogram of the total number of steps taken each day

```{r echo=TRUE,fig.width=10}
sum_steps <- activity[,sum(steps,na.rm=T),by=date]
library(ggplot2)
p <- ggplot(sum_steps,aes(date,V1))+geom_histogram(stat="identity")+ylab("total sum of steps taken per day")
print(p)
```

2.Calculate the mean and the median of the total number of steps taken per day

```{r}
mean_steps <- activity[,mean(steps,na.rm=T),by=date];
colnames(mean_steps) <- c("date","mean")
median_steps <- activity[,median(as.double(steps,na.rm=T)),by=date];
colnames(median_steps) <- c("date","median")
mean_median_steps <- merge(mean_steps,median_steps,by="date")
```

The mean of the total number of steps taken per day is showed in the following table:

```{r}
print(mean_median_steps)
```

#### Part 3: WHAT IS THE AVERAGE DAILY ACTIVITY PATTERN?

1.The average daily activity pattern

```{r fig.width=10}
ave_steps <- activity[,mean(steps,na.rm=T),by=interval]
plot(ave_steps,type="l",xlab="5-minute interval",ylab="average number of steps taken")
```

2.find 5-minute interval which contains the maximum number of steps:

```{r}
ind <- which.max(ave_steps$V1)
max_interval <- ave_steps$interval[ind]
```

The 5-minute interval on average across all the days in the dataset contains the maximum number of steps is `r max_interval`

#### Part 4: IMPUTING MISSING VALUES

1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
count <- sum(is.na(activity$steps))
```

The total number of missing values in the dataset is `r count`

2.Create a new dataset that is equal to the original dataset but with the missing data filled in. NOTE: using for loop to do this but I think there is a better way.

```{r}
new_activity <- activity;
for(i in 1:nrow(new_activity)){
        if(is.na(activity$steps)[i]==T){
                Ind2 <- which(ave_steps$interval==new_activity$interval[i]);
                new_activity$steps[i] <- ave_steps$V1[Ind2]
        }
}
```

3.make a histogram of steps taken each day of the new dataset

```{r fig.width=10}
sum_steps2 <- new_activity[,sum(steps),by=date]
p <- ggplot(sum_steps2,aes(date,V1))+geom_histogram(stat="identity")+ylab("total sum of steps taken per day")
print(p)
```

4.report the mean and median total number of steps taken per day of the new dataset

```{r}
s2 <- new_activity[,mean(steps),by=date]
median_step2 <- new_activity[,median(as.double(steps)),by=date]
```

as we can see, the values are differ from what we have seen in part one and the impact is: there is no NA any longer, which all be replaced by the average fo 5-interval minute

#### Part 5: ARE THERE DIFFERENCES IN ACTIVITY PATTERNS BETWEEN WEEKDAYS AND WEEKENDS?

1.Create a new factor variable in the dataset with two levels ¨C ¡°weekday¡± and ¡°weekend¡± indicating whether a given date is a weekday or weekend day.

```{r}
library(lubridate)
new_activity$weekdays <- wday(as.Date(new_activity$date))
Indweekday <- which(new_activity$weekdays!=6 & new_activity$weekdays!=7)
new_act_weekday <- new_activity[Indweekday,]
Indweekend <- which(new_activity$weekdays==6 | new_activity$weekdays==7)
new_act_weekend <- new_activity[Indweekend,]
```

2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r fig.height=4,fig.width=10}
par(mfrow=c(1,2))
ave_steps_weekday <- new_act_weekday[,mean(steps),by=interval]
plot(ave_steps_weekday,type="l",xlab="5-minute interval",ylab="average number of steps taken in weekday",main="weekday")

ave_steps_weekend <- new_act_weekend[,mean(steps),by=interval]
plot(ave_steps_weekend,type="l",xlab="5-minute interval",ylab="average number of steps taken in weekend",main="weekend")
```
From this two figures we can see that there are some sifferences in activity patterns between weekdays and weekends. 